Exploring the Impact of Queer Media Characters on LGBTQ+ Psychological Well-Being: The Role of Queer Media Consumption, Perceived Authenticity, and Intersectionality

### 2025-02-13 Manuscript.qmd
Update discussion and conclusion

Data visualization of bubble plots revealed that queer media exposure frequency and engagement may predict queer media characters’ psychological impacts on LGBTQ+ individuals. The bubble plot for queer media exposure and discussion frequency indicated that developed countries have more participants who engaged less with queer media characters, which may contribute to lower impact scores (Figure 7). Though bubbles for developing countries were smaller, it was the result of a much smaller sample size (N = 22) for developing countries' residents than for developed countries' residents (N = 145). The bubble plot featuring different gender identities also showed that women, nonbinary, and agender respondents have higher discussion frequency of queer media representations than men (Figure 8), which may contribute to higher community belong scores and the overall psychological impact scores.

@@

## Limitations and Future Directions

## Conclusion

@@

# Appendix

# Title for Appendix

>>>>

Data visualization of bubble plots revealed that queer media exposure frequency and engagement may predict queer media characters’ psychological impacts on LGBTQ+ individuals. The bubble plot for queer media exposure and discussion frequency indicated that developed countries have more participants who engaged less with queer media characters, which may contribute to lower impact scores (Figure 7). Though bubbles for developing countries were smaller, it was the result of a much smaller sample size (*n* = 22) for developing countries' residents than for developed countries' residents (*n* = 145). The bubble plot featuring different gender identities also showed that women, nonbinary, and agender respondents have higher discussion frequency of queer media representations than men (Figure 8), which may contribute to higher community belonging scores and the overall psychological impact scores.

@@

## Queer Media Characters' Psychological Impacts on Different LGBTQ+ Subgroups

The results of the study indicated that queer media characters' psychological impacts on LGBTQ+ individuals did not significantly differ among different generations. This finding suggests that queer media representations may have become more diverse in age over time, catering to different generations of LGBTQ+ individuals. It is also consistent with @dhoest_queer_2022, which found that LGBTQ+ individuals across all generations were keen to watch LGBTQ-themed films and rated TV shows featuring LGBTQ+ characters as important. The recent years' increase in transgender media representation may have contributed to the positive psychological impacts on transgender individuals, resulting in the insignificant differences between transgender and non-transgender individuals. For instance, the number of transgender characters on broadcast TV increased from two characters in 2014 to 24 characters in 2024 [@GLAAD2024]. Reality shows such as *RuPaul's Drag Race* and TV series like *Pose* and *Euphoria* were also popular among transgender participants in this study.

The study has also found that queer media characters' psychological impacts on LGBTQ+ individuals were less significant for individuals living in developed countries than for those living in developing countries. A plausible explanation could be a higher saturation of queer characters in media leading to a normalization of queer identities in developed countries. As a result, the impact of individual characters may be less pronounced, as audiences become desensitized to their presence. In contrast, the scarcity of queer characters in developing countries may lead to a more profound impact when they appear. Additionally, LGBTQ+ individuals in developed countries might have more access to mental health resources and support systems (e.g., LGBTQ+ advocacy groups), which could mitigate the effects of media representations on psychological well-being. Nevertheless, LGBTQ+ individuals in developing countries may rely more on cultural resources such as queer media characters for representation and community belonging, leading to a stronger psychological impact.

Regarding gender identity and racial identity, the study found that queer media characters' psychological impacts on LGBTQ+ individuals were more positive for gender and racial minorities. These findings reflected that queer media representations have become more diverse in gender and racial identities. Specifically, @GLAAD2024 reported 2024 as the fifth year there were more LGBTQ+ women than men on broadcast, with 58% were women or nonbinary, and 42% were men. Regarding racial identity, the number of LGBTQ+ characters of color on broadcast TV increased from 23% in 2014 to 50% in 2024, with 30% of LGBTQ+ characters being Black, 7% Latine, 5% Asian, and 8% multiracial or others. This shift in decades of LGBTQ+ representation may have contributed to the more positive psychological impacts on gender and racial minorities.

Nevertheless, the study found that regarding sexual orientation, the mainstream group (i.e., gay or lesbian) had the highest psychological impact scores, while bisexual respondents had the lowest. This is consistent with @GLAAD2024's report that the percentage of bisexual characters on broadcast TV was significantly lower than the percentage of bisexual as the majority in the LGBTQ+ community (24% vs. 58%). Apart from this disparity, the bi+ erasure (i.e., the tendency to not specify a character as bisexual) in media may have also contributed to the lower psychological impact scores on bisexual respondents. The following section further discusses the potential explanations from the perspectives of perceived character authenticity and intersectionality.

This study found that the frequency of queer media exposure and discussion engagement may predict psychological impacts. Specifically, the results indicated that more respondents in developed countries had lower exposure to queer media characters (e.g., a few times per year), which may contribute to lower psychological impact scores. This finding suggests that the normalization of queer characters and more well-established social support systems for LGBTQ+ individuals in developed countries may have led to a desensitization effect, reducing the psychological impacts of queer characters. Regarding queer media discussion engagement, the results revealed that women and nonbinary individuals had higher discussion engagement than men, which may contribute to higher community belonging scores and overall psychological impact scores. This finding also highlights that there may not be enough online forums or platforms specifically tailored to the needs and interests of queer men. Additionally, social norms and expectations around masculinity may discourage queer men from openly discussing their experiences or engaging in vulnerable conversations that reveal their identities. These norms often penalize emotional expressiveness in men, which might make them less likely to participate in discussions that require self-disclosure, resulting in lower community belonging scores.

The study also revealed that perceived queer character authenticity and intersectionality are strongly positively related to queer characters' psychological impact scores. The more authentic and intersectional LGBTQ+ individuals perceive queer media characters, the more positive psychological impacts they may experience. This is consistent with previous research that the amplification of stereotypes and lack of intersectional representation in media can lead to negative psychological outcomes for marginalized groups [@nolke_making_2018]. The findings suggest that the authenticity and intersectionality of queer media characters are crucial factors in determining their psychological impacts on LGBTQ+ individuals given the high explanatory power.

The correlations between authenticity and intersectionality ratings and psychological impact scores also explain the disparities in psychological impacts among different LGBTQ+ subgroups, especially for sexuality. Regarding gender identity, authenticity rating has a stronger effect on psychological impact scores for men than women and other gender identity. This finding suggests that queer men may be more sensitive to whether media representations feel authentic or genuine. This sensitivity could stem from past experiences with harmful stereotypes or depictions such as ‘perfect homo’ of queer men in media. The lower psychological impact scores for queer men could also result from more existing queer men stereotypes in the media than women and other gender identities. Additionally, queer men respondents were less sensitive to intersectionality rating since there was a significant proportion of the data points featuring high intersectionality with low psychological impact scores. This suggests that queer men might prioritize or respond more strongly to characters that focus primarily on their sexual orientation rather than on intersectional identities. Meanwhile, queer men who face fewer intersectional challenges may not feel represented by narratives that emphasize multiple forms of marginalization, leading to lower psychological benefit from such representations. Thus, the recent progress in queer media characters' intersectionality may result in more positive psychological impacts on women and other gender identities, as they were more invisible and marginalized before. This also applies to racial identity, where the intersectionality rating has a stronger effect on psychological impact scores for black, asian, and indian respondents than white respondents as they may have experienced more intersectional challenges in their lives. The higher psychological impact scores for racial minorities may also result from the recent increase in racial minority representation in queer media characters that feature more intersectional identities.

It is worth noting that authenticity and intersectionality ratings may also play a significant role in explaining the disparities in psychological impacts among different sexual orientations. Among all sexual orientations, only bisexual respondents had more data points condensed on the lower end of the authenticity rating and psychological impact scores. This suggests that bisexual respondents' low psychological impacts scores may be related to the perceived low authenticity of bisexual media characters. While bisexual men are often depicted as secretly gay with absence of bisexual representations, biseuxal women are often depicted as hypersexualized or fetishized [@young_why_2022]. This can further reinforce the stigma and stereotypes surrounding bisexuality as creedy or indecisive, perpetuating biphobia in society. The negative associations between intersectionality rating and psychological impact scores for asexual and pansexual men further indicated that queer men may prioritize or respond more strongly to characters that focus primarily on their sexual orientation rather than on intersectional identities as there was barely any asexual or pansexual men in the media.

## Conclusion
This study investigates the psychological impacts of queer media characters on LGBTQ+ individuals across different subgroups and identifies predictors of these impacts. The findings indicate that while queer media characters' psychological impacts do not significantly differ among different generations and transgender status, they are more positive for individuals living in developing countries, women and nonbinary individuals, gay or lesbian, and racial minorities, indicating a progress in queer media representations of more marginalized groups in recent years. The study also reveals that the authenticity and intersectionality of queer media characters are crucial factors in determining their psychological impacts on LGBTQ+ individuals. While queer men may be more sensitive to authenticity due to existing stereotypes, queer women and other identities tend to benefit more from intersectional representations. Additionally, bisexual individuals may experience lower psychological impacts due to the lack of authentic bisexual representations in media. The current study enriches the understanding of different LGBTQ+ individuals' needs for queer media characters by comparative analysis, revealing the necessity for the media industry to provide LGBTQ+ audience with more authentic and diversified queer images. It is also necessary to present bisexual characters in a non-stereotypical way to reduce biphobia and increase bisexual individuals' psychological well-being. The limitations of this study lie in the relatively small sample size, disproportionate representation of different subgroups by sample size, and lack of comparison among different media’s impacts. Future research may consider a larger sample with more balanced representation of different subgroups to further explore the psychological impacts of queer media characters on LGBTQ+ individuals. Scholars may also be interested in the intersectional analysis of different media’s impacts.

@@
